The Dynamics of Anthropomorphic Effects in Human-Robot Interaction

### 2014-03-21 paper.tex
Forum study part complete

We investigated the amount and context of anthropomorphic language in three online discussion forums about the robotic dog AIBO, the vacuum cleaning robot Roomba, and the tablet computer iPad. This analysis was done by means of a content analysis of 750 forum posts that were further partitioned in n = 1208 valid segments \citep{fink_anthropomorphic_2012}. These three devices, pictured in Figure \ref{fig:aibo-roomba-ipad}, were chosen as they show different designs and fulfill different purposes. We would describe (1) AIBO as a zoomorphic / pet-like robot that does not have a clear purpose but engages people to play with it; (2) Roomba as a mechanical / functional domestic service robot that has a practical main purpose (vacuum cleaning) but offers not much of interactivity; (3) the iPad as a minimalistic designed interactive (touch-based) technology with multiple purposes (both entertainment, and professional / practical use). Based on these different characteristics, we made the two following hypotheses concerning anthropomorphism:

@@

>>>>

We investigated the amount and context of anthropomorphic language in three online discussion forums about the robotic dog AIBO, the vacuum cleaning robot Roomba, and the tablet computer iPad. This was done by means of a content analysis of 750 forum posts that were further partitioned in n = 1208 valid segments \citep{fink_anthropomorphic_2012}. Hereafter, we refer to this study as the "Forum study". The three devices, pictured in Figure \ref{fig:aibo-roomba-ipad}, were chosen as they show different designs and fulfill different purposes. We would describe (1) AIBO as a zoomorphic / pet-like robot that does not have a clear purpose but engages people to play with it; (2) Roomba as a mechanical / functional domestic service robot that has a practical main purpose (vacuum cleaning) but offers not much of interactivity; (3) the iPad as a minimalistic designed interactive (touch-based) technology with multiple purposes (both entertainment, and professional / practical use). Based on these different characteristics, we made the two following hypotheses concerning anthropomorphism:

@@

Similar to how \cite{friedman_hardware_2003} analyzed anthropomorphic language in online forums, we coded a segment as anthropomorphic when the respective device was described in terms of:

\begin{itemize}
	\item \textit{Life-likeness}, uch as being alive or having (parts of) a body, \eg \textit{``she can be reborn since you copied her memory stick''};
	\item \textit{Emotional states} or the technology having a feeling, \eg \textit{``oh no, poor AIBO''};
	\item \textit{Gender, personality} or the technology having an \textit{intention}, \eg Roomba \textit{``seems to like to hang around under the sofa''};
	\item When the author gave the technology a \textit{name}, \eg \textit{``whenever I show Java his pink ball''};
	\item Being \textit{socially integrated}, such as being a family member, \eg \textit{``I'm considering adding a Roomba to the family''};
	\item \textit{Metaphorical ways}, \eg Roomba \textit{``sings its victory song when it finishes and docks''}.
\end{itemize}

In the next sections, we present findings in respect of our two hypotheses: anthropomorphic language related to the device, and anthropomorphic language related to the topic of the forum post.

Overall, our analysis revealed that in all three forums, discussions about \textit{technical} aspects of the respective device were most frequent (67.7~\% of all 1208 categorized segments). However there were huge differences between the three devices: \textit{technical} aspects made up 79.9~\% of the discussion in the Roomba forum, compared to 69.9~\% in the AIBO forum, and 55.2~\% in the iPad forum. Another fairly frequently discussed topic concerned the \textit{usage} of the respective device, thus people describing how and for what they use their technology. This topic made up 21.8~\% of all posts, and was most frequent in the iPad forum (35.7~\%), followed by the Roomba forum (15.9~\%), and the AIBO forum (10.7~\%). Last but not least, we defined \textit{relation} as the third topic of interest. Discussions in which people described their relation to the respective device made up 10.5~\% of all posts, and was most frequent in the AIBO forum (19.4~\%), followed by the iPad forum (9.1~\%), and the Roomba forum (4.2~\%). \hl{(if it is getting too much we can leave out this paragraph, it's not crucial, and doesn't say anything about anthropomorphism directly)}

\subsubsection{Anthropomorphic language related to the device (design)\\}

A statistical analysis revealed a significant difference for the use of anthropomorphism across the three devices \citep{fink_anthropomorphic_2012}. In accordance with our first hypothesis, the highest amount of anthropomorphic language (56.6~\%) was found in the AIBO forum. In contrast, only 11.7~\% and respectively 2.2~\% of the segments in the Roomba and the iPad forum contained anthropomorphic expressions.
The data suggests that on a subjective scale of perceived human-likeness (or rather life-likeness), Roomba is closer to the iPad than to the AIBO, at least in matters of how they are described with words. This was a bit against our expectation. Still, the Roomba is an autonomous robot that can navigate around the house itself, and we expected this property to enhance anthropomorphism. We interpret that here, the physical embodiment of the object and how far it resembles a life-like entity determine how far the object is anthropomorphized, and as such the robotic dog is anthropomorphized more than the functionally designed devices.

\subsubsection{Anthropomorphic language related to the topic of conversation\\}

Also in terms of anthropomorphic language related to the topic of conversation, a statistical analysis revealed a significant difference \citep{fink_anthropomorphic_2012}. The analysis suggests that anthropomorphizing happens more when the topic of conversation is \textit{relation}, compared to \textit{usage}, and \textit{technical} aspects. This finding is qualitatively confirmed when looking at the distribution of anthropomorphic expressions across the three topics in the Roomba and iPad forum (with both technologies the highest amount of anthropomorphism is in the \textit{relation} topic). However, in the AIBO the amount of anthropomorphic segments is highest for the \textit{usage} topic.
We interpret that it is the different type of interaction / use context that makes the difference. Contrary to the iPad and the Roomba, the AIBO is made to encourage its user in a playful / social interaction, which in turn is a typical use context. Furhter, AIBO has the ability to respond intelligently to its user and environment, which might also encourage people to ascribe human-like abilities to it while interacting with it.

\subsubsection{Summary Forum study\\}

Overall, findings of the content analysis suggest that there are differences in people's tendency to use anthropomorphic language, originating form the design characteristics of the artifact. Interestingly, not only the physical embodiment of the object but also the interaction complexity and purpose of use of the object seems to play a role in how far the object is anthropomorphized.
